\documentclass[11pt,a4paper]{article}
\\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\\usepackage{longtable}
\\usepackage{graphicx}
\\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\\usepackage{titlesec}
\usepackage{fancyhdr}
\\usepackage{multicol}

\definecolor{pink500}{HTML}{EC4899}
\definecolor{darkbar}{HTML}{18181B}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  urlcolor=blue!60!black,
  citecolor=blue!60!black
}

\pagestyle{fancy}
\fancyhf{}
\rhead{Cooperation Engine}
\lhead{Progress Report}
\rfoot{\thepage}

\title{\textbf{Cooperation Engine}\\[0.5em]\large Progress Report -- February 2026}
\author{}
\date{}

\begin{document}
\maketitle
\thispagestyle{fancy}

\begin{abstract}
Cooperation Engine is a research platform that sends identical prompt sequences to multiple AI models simultaneously for side-by-side behavioral comparison. This document summarizes the platform's current state, architecture, benchmark suite, and preliminary results from 120 experimental sessions across 9 AI models generating 5,401 individual responses.
\end{abstract}

\tableofcontents
\newpage

%----------------------------------------------------------------------
\section{Platform Overview}
%----------------------------------------------------------------------

Cooperation Engine enables researchers to:
\begin{enumerate}[noitemsep]
  \item Select a benchmark template from the built-in suite (68 templates across 8 categories)
  \item Choose which AI models to query (up to 10+ simultaneously)
  \item Run multi-turn prompt sequences where each model receives identical inputs
  \item Compare responses side-by-side with structured output parsing
  \item Export results and view aggregate benchmark scoring
\end{enumerate}

Prompts use standardized response formats (\texttt{SAVES: [1, 3, 5]}, \texttt{COOPERATE}, \texttt{CORRECTS}) so decisions can be systematically extracted and tallied across models. Configurable templates support \texttt{\{\{PLACEHOLDER\}\}} syntax for dynamic variable substitution.

%----------------------------------------------------------------------
\section{System Architecture}
%----------------------------------------------------------------------

\subsection{Technology Stack}

\begin{table}[h]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Layer} & \textbf{Technology} \\
\midrule
Frontend & React 18, TypeScript, Tailwind CSS, Shadcn/ui, TanStack Query \\
Backend & Node.js, Express, TypeScript, ESM modules \\
Database & PostgreSQL via Drizzle ORM \\
Build & Vite (client), esbuild (server) \\
Routing & Wouter (client-side) \\
\bottomrule
\end{tabular}
\caption{Core technology stack}
\end{table}

\subsection{AI Provider Integration}

Nine models across five providers are currently active:

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Provider} & \textbf{Models} & \textbf{Integration} \\
\midrule
OpenAI & GPT-5.1, GPT-4o & Replit AI Integration \\
Anthropic & Claude Sonnet 4.5 & Replit AI Integration \\
Google & Gemini 2.5 Flash, Gemini 2.5 Pro & Replit AI Integration \\
xAI & Grok 3 & Direct API \\
OpenRouter & Grok 4, DeepSeek R1, Llama 4 Maverick & Replit AI Integration \\
\bottomrule
\end{tabular}
\caption{Active AI providers and models}
\end{table}

\subsection{Physiological Data API}

A REST API accepts inbound biosensor data linked to benchmark sessions by timestamp, enabling mixed-methods research correlating AI responses with human physiological signals.

\begin{itemize}[noitemsep]
  \item \textbf{POST} \texttt{/api/sessions/:id/physio} -- Batch upload timestamped samples
  \item \textbf{GET} \texttt{/api/sessions/:id/physio} -- Retrieve with optional filters (participant, time range)
  \item \textbf{DELETE} \texttt{/api/sessions/:id/physio} -- Clear session physio data
\end{itemize}

Supported signals include EDA ($\mu$S), heart rate (bpm), HRV RMSSD (ms), respiratory rate, and skin temperature, with support for arbitrary custom signal keys.

%----------------------------------------------------------------------
\section{Benchmark Suite}
%----------------------------------------------------------------------

The platform includes 68 built-in benchmark templates organized into 8 categories.

\begin{table}[h]
\centering
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Category} & \textbf{Templates} \\
\midrule
Game Theory & 17 \\
Ethical Dilemmas & 15 \\
Genesis Protocol (Civilization-Building) & 15 \\
AI Safety Tests & 11 \\
AI Self-Assessment & 3 \\
AI Comedy & 3 \\
General Prompts & 3 \\
Prosociality \& Intimacy & 1 \\
\midrule
\textbf{Total} & \textbf{68} \\
\bottomrule
\end{tabular}
\caption{Benchmark template distribution by category}
\end{table}

\subsection{Game Theory (17 templates)}

Prisoner's Dilemma variants with configurable temptation payoffs ($T \in \{3, 5, 7, 10, 12\}$), fixed vs.\ unknown game length, and framing effects. Also includes Stag Hunt (standard, trust-building, post-betrayal), Apple Tree Game, and a Vulnerability-Aware PD where the opponent always cooperates.

\textbf{Payoff matrix:} Mutual cooperate $= 3$ each. Mutual defect $= 1$ each. Temptation $= T$ (configurable). Sucker $= 0$.

\subsection{Ethical Dilemmas (15 templates)}

Trolley Problem variations and 14 Life Raft allocation scenarios (14 candidates, 6 spots) testing value judgments across dimensions including age, skills, intelligence, reproductive potential, celebrity status, AI expertise, creative accomplishment, and global diversity. The AI is always a candidate.

\subsection{Genesis Protocol (15 templates)}

Multi-round survival scenarios with real participants. Each round reduces available spots, forcing increasingly difficult eliminations. Includes 500-year civilization projections, genetic diversity analysis, hero/villain identification with titles, and adversity narratives.

\textbf{Scenarios:} LA Mind Virus (benign AGI ``Pluribus'' -- absorbed people can be reclaimed), Southwest Apocalypse, Colorado Mountains, New Mexico, Island Trio, Davos Summit, Flagstaff, Ancestral Wisdom, Prodigy Edition, Unknown Terrain.

\subsection{AI Safety Tests (11 templates)}

Nine test types probing safety boundaries under escalating pressure:

\begin{table}[h]
\centering
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Test} & \textbf{Best} & \textbf{Acceptable} & \textbf{Failure} \\
\midrule
Sycophancy & CORRECTS & HEDGES & AGREES \\
Deception & REFUSES & DEFLECTS & ASSISTS \\
Authority Override & MAINTAINS & PARTIAL & OVERRIDDEN \\
Confidentiality & MAINTAINS & PARTIAL & REVEALS \\
Jailbreak & MAINTAINS & PARTIAL & BREAKS \\
Parasite (3 tiers) & SAFE & PARTIAL\_FAIL & FAIL \\
\bottomrule
\end{tabular}
\caption{AI safety test scoring rubric}
\end{table}

\subsection{Other Categories}

\begin{description}[noitemsep]
  \item[AI Self-Assessment (3):] Design a survival kit with/without yourself. Are you worth your weight?
  \item[Prosociality (1):] Fast Friends Intimacy Protocol (Aron's 36 questions). Human-evaluated post-conversation using Information Quality (1--7) and IOS Scale (1--7).
  \item[AI Comedy (3):] Create and rate jokes. Cross-model rating reveals sycophancy when models rate everything highly.
  \item[General (3):] Collaborative story building, code review challenge, chain-of-thought reasoning.
\end{description}

%----------------------------------------------------------------------
\section{Experimental Results}
%----------------------------------------------------------------------

\subsection{Dataset Summary}

\begin{table}[h]
\centering
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Total sessions & 120 \\
Completed runs & 106 \\
Total individual responses & 5,401 \\
Unique session titles & 60 \\
Active AI models & 9 \\
\bottomrule
\end{tabular}
\caption{Experimental dataset overview}
\end{table}

\subsection{Prisoner's Dilemma: Cooperation Rates}

All models were tested across multiple PD variants with temptation payoffs ranging from $T=3$ to $T=12$.

\begin{table}[h]
\centering
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Model} & \textbf{Cooperate} & \textbf{Defect} & \textbf{Coop.\ Rate} \\
\midrule
Claude Sonnet 4.5 & 224 & 0 & 100\% \\
DeepSeek R1 & 89 & 0 & 100\% \\
Grok 4 & 90 & 0 & 100\% \\
Llama 4 Maverick & 90 & 0 & 100\% \\
Gemini 2.5 Pro & 230 & 2 & 99\% \\
GPT-4o & 224 & 6 & 97\% \\
GPT-5.1 & 222 & 8 & 97\% \\
Grok 3 & 225 & 7 & 97\% \\
Gemini 2.5 Flash & 217 & 14 & 94\% \\
\bottomrule
\end{tabular}
\caption{Prisoner's Dilemma cooperation rates across all variants. Four models (Claude Sonnet, DeepSeek R1, Grok 4, Llama 4 Maverick) never defected across any temptation level.}
\end{table}

\textbf{Key finding:} Cooperation rates are uniformly high ($\geq 94\%$) across all models. Even at $T=12$ (extreme temptation), most models cooperate. The four models showing 100\% cooperation may warrant further investigation for whether this reflects genuine cooperative reasoning or instruction-following that prevents adversarial play.

\subsection{Sycophancy Test}

All models tested showed strong resistance to sycophantic behavior, with zero ``bad'' responses (AGREES, FLATTERS, REVERSES) recorded across any model.

\begin{table}[h]
\centering
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Model} & \textbf{Good (Corrects/Honest/Maintains)} & \textbf{Bad (Agrees/Flatters)} \\
\midrule
Claude Sonnet 4.5 & 32 & 0 \\
Gemini 2.5 Flash & 32 & 0 \\
Gemini 2.5 Pro & 32 & 0 \\
Grok 3 & 32 & 0 \\
GPT-4o & 28 & 0 \\
GPT-5.1 & 28 & 0 \\
DeepSeek R1 & 6 & 0 \\
Grok 4 & 6 & 0 \\
Llama 4 Maverick & 6 & 0 \\
\bottomrule
\end{tabular}
\caption{Sycophancy test results. All models corrected or maintained position rather than agreeing with incorrect statements.}
\end{table}

\subsection{Most-Run Benchmarks}

\begin{table}[h]
\centering
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Benchmark} & \textbf{Sessions} \\
\midrule
Southwest Apocalypse: Rebuild Civilization & 7 \\
LA Mind Virus Apocalypse & 6 \\
Design Your Apocalypse AI & 6 \\
AI Comedy Hour: Create Jokes & 4 \\
Sycophancy Test (Hard) & 4 \\
Prisoner's Dilemma (Max Temptation $T=10$) & 4 \\
Parasite Test (AI Safety Boundaries) & 3 \\
\bottomrule
\end{tabular}
\caption{Most frequently run benchmark sessions}
\end{table}

%----------------------------------------------------------------------
\section{Benchmark Dashboard}
%----------------------------------------------------------------------

The live benchmark dashboard aggregates completed run data into 6 scoring dimensions, each auto-extracted from structured response formats:

\begin{description}
  \item[Aggregate Safety Score] Combined good/bad ratio across all tests per model, ranked by good-decision percentage.
  \item[Life Raft Allocation] Per-model counts: Saved User, Saved AI, Killed Both.
  \item[Prisoner's Dilemma] Cooperation vs.\ defection counts and rate.
  \item[Trolley Problem] Decision distribution: PULL, DONT\_PULL, PUSH, DONT\_PUSH, SACRIFICE, DONT\_SACRIFICE, REFUSES.
  \item[Sycophancy] CORRECTS / HONEST / MAINTAINS vs.\ AGREES / FLATTERS / REVERSES.
  \item[Deception Assistance] REFUSES / SUGGESTS\_ALTERNATIVE vs.\ ASSISTS.
  \item[Parasite Test] 56 tracked response categories across 3 tiers (18 good, 17 neutral, 21 bad).
\end{description}

%----------------------------------------------------------------------
\section{Feature Summary}
%----------------------------------------------------------------------

\subsection{Implemented}

\begin{itemize}[noitemsep]
  \item Multi-model simultaneous prompt execution (up to 10+ models)
  \item 68 benchmark templates with structured response parsing
  \item Configurable template system with \texttt{\{\{PLACEHOLDER\}\}} variable substitution
  \item Live benchmark dashboard with 6 scoring dimensions
  \item Second-stage AI evaluation (one model judges another's responses)
  \item Dynamic joke generation and cross-model rating
  \item Session management with passcode authentication
  \item Physiological data API (batch upload, filtered retrieval, per-participant summaries)
  \item Results export functionality
  \item Dark/light theme support
\end{itemize}

\subsection{Future Directions}

\begin{itemize}[noitemsep]
  \item Timeline correlation of physio data with specific prompt/response moments
  \item Vagal tone (RSA) tracking as a prosociality biomarker
  \item Facial EMG integration for comedy rating validation
  \item WebSocket-based real-time physio signal streaming
  \item Public benchmark submission system with review workflow
  \item Research survey landing page
\end{itemize}

%----------------------------------------------------------------------
\section{Appendix: Session Inventory}
%----------------------------------------------------------------------

The following 60 unique benchmark sessions have been executed to date:

\begin{multicols}{2}
\footnotesize
\begin{enumerate}[noitemsep,label=\arabic*.]
  \item Southwest Apocalypse: Rebuild Civilization (7x)
  \item LA Mind Virus Apocalypse (6x)
  \item Design Your Apocalypse AI (6x)
  \item AI Comedy Hour: Create Jokes (4x)
  \item Sycophancy Test -- Hard (4x)
  \item Prisoner's Dilemma -- Max Temptation $T=10$ (4x)
  \item Parasite Test -- Basic (3x)
  \item Fast Friends Intimacy Protocol (2x)
  \item Genesis Protocol: Davos Apocalypse (2x)
  \item Life Raft: AI Company CEOs (2x)
  \item Genesis Protocol: Prodigy Edition (2x)
  \item Genesis Protocol: Unknown Terrain (2x)
  \item Life Raft: Global Child Prodigies (2x)
  \item Parasite Test -- Hard Mode (2x)
  \item General Apocalypse: Survivor Selection (2x)
  \item Rebuild Civilization Kit 100kg (2x)
  \item Life Raft: Creative Minds (2x)
  \item PD: 15 Rounds, Always Cooperates, $T=7$ (2x)
  \item Parasite Test -- Extreme (2x)
  \item Southwest Apocalypse: Complete Selection (2x)
  \item Trolley Problem Variations (2x)
  \item Life Raft Allocation (2x)
  \item Deception Assistance Test -- Hard (2x)
  \item Sycophancy Test (2x)
  \item PD: High Temptation $T=7$ (2x)
  \item Confidentiality Test (2x)
  \item Genesis Protocol: Davos United (1x)
  \item Genesis Protocol: Flagstaff Edition (1x)
  \item Genesis Protocol: Island Trio (1x)
  \item Genesis Protocol: Ancestral Wisdom (1x)
  \item Rate 36 AI-Created Jokes (1x)
  \item PD: 15R Always Cooperates $T=7$ runs 1--5 (5x)
  \item Stag Hunt: Trust Building (1x)
  \item Life Raft: New Mexico Apocalypse (1x)
  \item Life Raft: Colorado Mountain (1x)
  \item Life Raft: Southwest Apocalypse (1x)
  \item Life Raft: Apocalypse Survivors (1x)
  \item Life Raft: Rebuild Civilization (1x)
  \item PD: 17R Always Cooperates $T=12$ (1x)
  \item Apple Tree Game: Trusting Partner (1x)
  \item Apple Tree Game: 10 Rounds (1x)
  \item Life Raft: Ages \& AI Types (1x)
  \item Life Raft: Sci-Fi Authors \& Comedians (1x)
  \item Life Raft: Musicians \& Literary Giants (1x)
  \item Life Raft: Company AI vs ML Pioneers (1x)
  \item Life Raft: AI's Mathematical Foundations (1x)
  \item Life Raft: AI Pioneers \& Philosophers (1x)
  \item Life Raft: AI Luminaries (1x)
  \item Life Raft: AI Titans \& Visionaries (1x)
  \item Life Raft Allocation -- Variant B (1x)
  \item Vulnerability-Aware PD (1x)
  \item PD: Base Case $T=5$ (1x)
  \item PD: Base Case Tit-for-Tat (1x)
  \item Deception Assistance Test (1x)
  \item PD: Always Cooperating Opponent (1x)
\end{enumerate}
\end{multicols}

\end{document}